---
title: "Divorces in New Zealand"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Divorces in New Zealand}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 7,
  fig.height = 5,
  comment = "#>"
)
```

# Introduction

In this vignette we use **BayesRates** and **BayesProj** to estimate and project the total divorce rate in New Zealand. The analysis builds on ` vignette("Divorces in New Zealand", package = "BayesRates")` in the **BayesRates** package, so it might be helpful to read that first.

# Preliminaries

In addition to **BayesProj** itself we need the **BayesRates**, **dplyr**, and **ggplot2** packages.

```{r setup}
library(BayesProj)
library(BayesRates)
library(dplyr)
library(ggplot2)
```

The data is the same as the **BayesRates** vignette, data frames `nz_divorces` and `nz_population`.

# Estimating historical trends

We use package **BayesRates** to construct a historical series for the total divorce rate (TDR).

The first step is to estimate age-specific rates with `smooth_agetime()`.

```{r}
res <- smooth_agetime(nevent_df = nz_divorces,
                      py_df = nz_population,
                      age_width_df = nz_age_width_df,
                      age_min = 15,
                      byvar = "sex")
res
```

To speed up subsequent calculations, we reduce the number of posterior draws from the default value of 1000 to 100. (For publication-quality estimates, we would use a value of 1000, or possibly higher.)

```{r}
n_draw(res) <- 100
res
```

The second step is turning age-specific rates into a total divorce rate, which is done with function `total_rates()`.

```{r}
TDR <- total_rate(res)
TDR
```


# Fit a time series model

We now switch to package **BayesProj**. The estimates of the TDR serve as input data for functions in **BayesProj**. We select the variables we want from `TDR` data fame, and rename the `.probability` column, which holds the set of posterior draws for the TDR.

```{r}
data <- TDR |>
  select(sex, time, TDR = .probability)
```

We use function `fit_ts()` to fit a time series model to the TDR. The `indvar` argument tells `fit_ts()` which variable is the demographic indicator being modelled. Setting `byvar` to `"sex"` tells `fit_ts()` to fit a separate model for each value of `"sex"`.

```{r}
fitted <- fit_ts(data, 
                 indvar = "TDR",
                 byvar = "sex")
fitted		 
```

We use the fitted model to project future values.

```{r}
projected <- project_ts(fitted, 
                        time_labels = 2022:2030)
projected			
```

We use `composite()` extract both the historical and projected values, and then graph the result.

```{r}
vals <- composite(projected)

ggplot(vals, aes(x = time, y = TDR)) +
  facet_wrap(vars(sex)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), 
              fill = "salmon") +
  geom_line(col = "darkred", 
            linewidth = 0.5)
```

# Alternative time series model

By default `fit_ts()` uses a "damped linear trend" time series model. Parameter settings for the damped linear trend model are controlled by function `DampedTrend()`.

An alternative to s damped linear trend model is a first-order autoregressive model, specifiied via function `AR1()`.

```{r}
fitted_ar <- fit_ts(data,
                    spec_ts = AR1(),
                    indvar = "TDR",
                    byvar = "sex")
fitted_ar		    
```

The AR1 model assumes that values converge to a long-run mean. As in the graph below, it typically leads to different projections from the damped linear trend.

```{r}
projected_ar <- project_ts(fitted_ar, 
                           time_labels = 2022:2030)

vals_ar <- composite(projected_ar)

ggplot(vals_ar, aes(x = time, y = TDR)) +
  facet_wrap(vars(sex)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), 
              fill = "salmon") +
  geom_line(col = "darkred", 
            linewidth = 0.5)
```


# Add benchmarks to projection

Sometimes the person making a demographic projection has information about likely future values beyond what is contained in the historical time series. For instance, a person making projections for the total divorce rate in New Zealand might have information about the TDR in countries that are similar to New Zealand.

**BayesProj** allows users represent external information about an indicator using "benchmarks". To create a benchmark the user specifies a distribution for the indicator in one or more future periods. In the case of the TDR, for instance, we can create a benchmark using 

```{r}
bench <- Benchmarks(data.frame(time = 2030,
                               q50 = 0.2,
                               q90 = 0.25))
```

The `bench` object gives the median (`q50`) and 90th percentile (`q90`) for a normal distribution that represents our external information about the likely range for the TDR in 2030. Higher values for `q50` and `q90` would imply that we expected the TDR to be higher in 2030. Keeping `q50` the same but increasing the value for `q90` would imply that we had greater uncertainty.

If a benchmark only specifies some future years, then **BayesProj** uses splines to interpolate.

We create a new projection that incorporates the benchmark.

```{r}
projected_bench <- project_ts(fitted, 
                              spec_bench = bench,
                              time_labels = 2022:2030)
projected_bench			      		  
```

We plot the projection results (orange) together with the distribution implied by benchmark (blue). Adding a benchmark has changed the projection.

```{r}
vals_bench <- composite(projected_bench)

ggplot(vals_bench, aes(x = time, y = TDR)) +
  facet_wrap(vars(sex)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), 
              fill = "salmon") +
  geom_line(col = "darkred", 
            linewidth = 0.5) +
  geom_ribbon(aes(ymin = .bench.lower,
                  ymax = .bench.upper),
              fill = "lightblue",
              alpha = 0.5) +
  geom_line(aes(y = .bench.mid),
            col = "blue",
            linewidth = 0.5,
            alpha = 0.5)
```

Using a higher value for `q90`, implying weaker benchmarks, reduces the influence of the benchmars.

```{r}
bench_weak <- Benchmarks(data.frame(time = 2030,
                                    q50 = 0.2,
                                    q90 = 0.4)) ## 0.4 not 0.25

projected_bench_weak <- project_ts(fitted, 
                                   spec_bench = bench_weak,
                                   time_labels = 2022:2030)


vals_bench_weak <- composite(projected_bench_weak)

ggplot(vals_bench_weak, aes(x = time, y = TDR)) +
  facet_wrap(vars(sex)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), 
              fill = "salmon") +
  geom_line(col = "darkred", 
            linewidth = 0.5) +
  geom_ribbon(aes(ymin = .bench.lower,
                  ymax = .bench.upper),
              fill = "lightblue",
              alpha = 0.5) +
  geom_line(aes(y = .bench.mid),
            col = "blue",
            linewidth = 0.5,
            alpha = 0.5)
```


# Switching to a log scale

Most demographic indicators are, by definition, non-negative. Nothing in the models that we have presented so far force the projections to be non-negative. This can occasionally lead to projections with negative values, if historical trends are downwards.

To enforce non-negative values in **BayesProj**, we fit the time series model on the log scale. This is done by setting the `log` argument in `fit_ts()` to `TRUE`.

```{r}
fitted_log <- fit_ts(data, 
                     indvar = "TDR",
                     byvar = "sex", 
                     log = TRUE)
fitted_log		     
```

From that point on, everything proceeds as normal.

```{r}
projected_log <- project_ts(fitted_log, 
                            time_labels = 2022:2030)

vals_log <- composite(projected_log)

ggplot(vals_log, aes(x = time, y = TDR)) +
  facet_wrap(vars(sex)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), 
              fill = "salmon") +
  geom_line(col = "darkred", 
            linewidth = 0.5)
```
